comgen-outline

### 2017-09-01 paper/comgen-paper.Rmd
edits to model section; caption on simulations figure

If generalization from observations can be described using probability, then it seems very natural to consider the semantic content of an utterance that conveys generalization also to be relating to probability. 
One very simple semantic theory that has been used for describing quantified statements (e.g., "some", "all") is a *threshold semantics*.
For generalizations, the threshold would apply to the subjective probability or *prevalence* $h$:

\begin{eqnarray}
\delta_{\denote{u_{gen}}(h, \theta)} & = & P(u_{gen} \mid h, \theta) \\
&\propto  &
\begin{cases}
1 & \text{if } h > \theta \\
0 & \text{otherwise}
\end{cases}
\end{eqnarray}

<!-- Generics express a relation between a kind K (e.g., \textsc{dogs}) and a property F (e.g., \textsc{is friendly}), such that the property can also be said to be applicable to an individual (i.e., "My dog is friendly").  -->
<!-- Bare plural statements (e.g., "Dogs are friendly") tend to yield a generic meaning, though other forms can express such a meaning [e.g., "A mongoose eats snakes."; @Carlson1977] . -->
<!-- Given that generics express a property that can be applied to individuals, it seems intuitive that what makes the statement true or false would correspond to the number of individuals who hold the property. -->
<!-- At first glance, generics convey that *most* have the property, such as "Most dogs are friendly". -->
<!-- "Robins lay eggs" is true, though, despite the fact that only female robins lay eggs. -->
<!-- If generic statements can be true by restricting themselves to only be describing females, why does *Robins are female* seem like a weird statement to make? -->
<!-- Finally, "Mosquitos carry malaria" is intuitively true despite malaria only being present in a tiny fraction of a percentage of mosquitos. -->
<!-- It seems that any hard condition on the prevalence of the feature in the category (e.g., how many robins lay eggs) would violate intuitions.  -->

<!-- In the case of generic language, we posit that conceptual factors may influence the meaning of a generic statement via its influence on subjective probability. -->
<!-- If that's true, a simple, semantic theory  -->

$\delta_{\denote{u}(h, \theta)}$ is the Kronecker delta function returning $1$ for probabilities $h$ greater than threshold $\theta$ and $0$ otherwise. 
Logical quantifiers can be described using threshold semantics (i.e., *some* is $h > 0$, all is $h = 1$).
What threshold $\theta$ should be used for the generic generalization?

To answer this, we draw a metaphor to Bayesian statistics and data analysis.
When there is a latent parameter the Bayesian scientist does not know much about, she puts a prior distribution over the parameter and infers the parameter's likely value by collecting data.
In a similar vein, we posit that the meaning of a generalization in language is *underspecified*; that is, there is not a fixed $\theta$ that corresponds to the generalization, but rather the listener has uncertainty over $\theta$ (in the form of a prior distribution $P(\theta)$) and infers $\theta$ in context (i.e., after having heard the utterance).
Though it is possible different $P(\theta)$ could be learned for different contexts, we assume this prior is identical across contexts and is completely uninformed by the context: $P(\theta) = \text{Uniform}(0, 1)$.

@@

Equation \ref{eq:L0} is a model of a Bayesian listener $L$ who updates her beliefs about the prevalence $h$ according the meaning of an utterance $u$.
$\delta_{\denote{u}(h, \theta)}$ is the likelihood term in Bayes' Rule and represents the literal semantic content of an utterance.
$P(h)$ represents the listener's prior world knowledge concerning the relevant prevalences $h$.
The particular form of the prior plausibly depends upon the kind of property or event in question and could be structured as a result of deeper conceptual knowledge about properties and events [@Griffiths2005; @Tenenbaum2006; @Kemp2008].
This component of the model provides a potential bridge to conceptual account of generics [e.g., @Leslie2007; @Prasada2006; @Prasada2013].
Critically in our model, this structure is in listener's knowledge of the world $P(h)$ and not in the semantic content of the utterance $u$. 
<!-- Our proposal that generic language and other forms of communicating generalizations convey information about probabilities that underly generalization and prediction implies that the conceptual factors of the kind discussed by Prasada, Leslie, et al. influence the meaning of a generic statement via their influence on subjective probability. -->
<!-- Flexible, structured models of the world have been shown to do just that in the case of generalizations from observations, merging structured representations with probabilisitic inference [e.g., @Kemp2008; @Goodmanconcepts]. -->
We term this prior distribution the **prevalence prior**.

@@

This is a Bayesian model of interpreting a statement conveying a generalization whose meaning is simple (i.e., a threshold) but underspecified.
Figure \ref{fig:simulations} shows model $L$'s posterior distribution on prevalence $h$ upon hearing utterances corresponding to the uncertain threshold (Eq. \ref{eq:L0}) as well as a comparison model using a fixed threshold at 0 (intuitively, corresponding to the quantifier "some").
The differences in the semantics are easily seen when considering a uniform prior distribution over $h$ (top left facet).
The fixed threshold model rules out only the lowest possible prevalence ($h=0$) and returns a posterior very similar to the prior.
The behavior of the generalization is distinct: The higher the $h$, the more likely it is after hearing the generalization.

The intuition behind the model's behavior can be gleaned by imagining what the listener $L$ would believe given different thresholds and averaging over those possibilities.
If the threshold $\theta$ is very high, only the highest prevalence $h$ would pass the threshold.
If the threshold $\theta$ is slightly lower but still high, the highest $h$ would still pass the threshold, as would some that are slightly lower.
As the threshold $\theta$  takes on lower and lower values, more and more $h$'s would surpass it.
When $\theta$ is very low, almost all $h$ are consistent with it (akin to fixed-threshold model).
When the listener averages over these possibilities, $h$'s with higher values pass more thresholds; thus, higher $h$'s are more likely *a posteriori*.

In this paper, we are interested in explaining the variable *endorsements* of generalizations (e.g., "Birds lay eggs" is good; "Birds are female" is not) based on some known or believed prevalence $h$ (e.g., only female birds lay eggs).
Following the Rational Speech Act (RSA) framework [@Goodman2016], we describe an endorsement model as a Bayesian decision of whether or not to say the generalization to the listener $L$ (Eq. \ref{eq:L0}).^[A more general version of this model can relax the assumption that the speaker has access to a specific value of prevalence $h$ that she is trying to communicate.
Rather, the speaker may have a belief distribution over $h$ for the category $k$, corresponding to the speaker's beliefs about the prevalence of the property in the category. 
In this situation, we would define the speaker's utility to be the expected value of the informativity, which integrates over her belief distribution:
$S(u \mid k) \propto \exp{(\lambda \cdot {\mathbb E}_{h\sim P_{k}} \ln{ \int_{\theta} L(h, \theta \mid u)} \diff \theta )}$
]

\begin{equation} 
S(u \mid h) \propto \exp{(\lambda \cdot \ln{ \int_{\theta} L_0(h, \theta \mid u)}  \diff \theta  }
\label{eq:S1}
\end{equation}

In RSA, language production and comprehension are defined by recursive Bayesian models describing speaker and listener.
Following RSA, the speaker is assumed to be an approximately rational Bayesian agent with degree with rationality governed by parameter $\lambda$.
Since the speaker $S$ doesn't know the threshold $\theta$, she must integrate over the likely values that the listener will infer.

@@

```{r simulations, fig.width = 9, fig.height=5}
all.simulations %>%
  mutate(Distribution = paste(Distribution, utterance, sep = "_")) %>%
  filter(Distribution != "Some", Distribution != "Most", Distribution != "Prior_generic") %>%
  mutate(Distribution = factor(Distribution, levels = 
                             c("Prior_some",  "Posterior_Generalization_generic", "Posterior_Generalization_some"),
                                labels = c("Prior/Silence",  "Generalization (Uncertain threshold)", "Fixed threshold")),
         PriorShape = factor(PriorShape, levels = c("uniform", "barks","laysEggs","carriesMalaria","isFemale"),
                             labels = c("uniform", "bark","lay eggs","carry malaria","are female")
                                  #"eatsFood","areSick", "uniformRare", 
                                  #"carriesMalaria"
                                  )
         ) %>%
  ggplot(., aes(x = value, #y = prob, 
                #lty = Distribution, 
                color = Distribution, fill = Distribution))+
    geom_density(#stat= 'identity', 
                aes(y = ..scaled..),
                 size = 0.6, alpha = 0.7, adjust = 1.5)+
    #geom_bar(stat = 'identity', position = position_dodge(), color = 'black')+
    theme_few() +
    #scale_color_solarized()+
    #scale_fill_solarized()+
    #scale_fill_manual(values = c("#636363", "#abdda4", "#2b83ba", "#d7191c"))+
    #scale_color_manual(values = c("#636363", "#abdda4", "#2b83ba", "#d7191c"))+
   scale_fill_manual(values = c("#636363", "#d7191c", "#2b83ba"))+
    scale_color_manual(values = c("#636363", "#d7191c", "#2b83ba"))+
    #scale_linetype_manual(values = c(3, 4, 2, 1))+
    scale_linetype_manual(values = c(3, 4, 1))+
    scale_x_continuous(breaks = c(0, 0.5, 1), limits= c(0, 1))+
    scale_y_continuous(breaks = c(0, 1), limits= c(0, 1))+
    xlab("Degree of probability h") +
    ylab("Scaled probability density") +
  facet_grid(PriorShape~Distribution, scales = 'free')+
  guides(fill = F, color = F)+
  theme(strip.text.y = element_text(angle = 0))
```

>>>>

If generalization from observations can be described using probability, then it seems  natural to consider the semantic content of an utterance that conveys generalization also to be relating to probability. 
One very simple semantic theory that has been used for describing quantified statements (e.g., "some", "all") is a *threshold semantics*.
For generalizations, the threshold would apply to the subjective probability or *prevalence* $h$:

\begin{eqnarray}
\delta_{\denote{u_{gen}}(h, \theta)} & = & P(u_{gen} \mid h, \theta) \label{eq:delta} \\ 
&\propto  &
\begin{cases}
1 & \text{if } h > \theta \\
0 & \text{otherwise}
\end{cases}
\end{eqnarray}

$\delta_{\denote{u}(h, \theta)}$ is the Kronecker delta function that returns probabilities proportional to $1$ for prevalence values $h$ greater than threshold $\theta$, and $0$ otherwise. 
Logical quantifiers can be described using threshold semantics (i.e., *some* is $h > 0$, all is $h = 1$).
What threshold $\theta$ should be used for the generic generalization?

To answer this, we draw a metaphor to Bayesian statistics and data analysis.
When there is a latent parameter the Bayesian scientist knows little about, she puts a prior distribution over the parameter and infers the parameter's likely value by collecting data.
In a similar vein, we posit that the meaning of a generalization in language is *underspecified*; that is, there is not a fixed $\theta$ that corresponds to the generalization, but rather the listener has uncertainty over $\theta$---in the form of a prior distribution $P(\theta)$---and infers $\theta$ after hearing an utterance.
Though it is possible different $P(\theta)$ could be learned for different contexts, we assume this prior is identical across contexts and is completely uninformed by the context: $P(\theta) = \text{Uniform}(0, 1)$.

Given the likelihood "meaning" function defined in Eq. \ref{eq:delta} and the prior distribution over thresholds $\theta$, we can formalize a **listener model** by specifying a prior distribution over the prevalence $P(h)$:

@@

Equation \ref{eq:L0} is a model of a Bayesian listener $L$ who updates her beliefs about the prevalence $h$ according the meaning of an utterance $u$.
$\delta_{\denote{u}(h, \theta)}$ is the likelihood term in Bayes' Rule and represents the literal semantic content of an utterance.
$P(h)$ represents the listener's prior world knowledge concerning the relevant prevalences $h$.
The particular form of the **prevalence prior** $P(h)$ plausibly depends upon the kind of property or event in question and could be structured as a result of deeper conceptual knowledge about properties and events [@Griffiths2005; @Tenenbaum2006; @Kemp2008].
This component of the model provides a potential bridge to conceptual account of generics [e.g., @Leslie2007; @Prasada2006; @Prasada2013].
Critically in our model, this structure is in listener's knowledge of the world $P(h)$ and not in the semantic content of the utterance $u$.

<!-- Our proposal that generic language and other forms of communicating generalizations convey information about probabilities that underly generalization and prediction implies that the conceptual factors of the kind discussed by Prasada, Leslie, et al. influence the meaning of a generic statement via their influence on subjective probability. -->
<!-- Flexible, structured models of the world have been shown to do just that in the case of generalizations from observations, merging structured representations with probabilisitic inference [e.g., @Kemp2008; @Goodmanconcepts]. -->

@@

Figure\ \@ref(fig:simulations) shows the listener model $L$'s posterior distribution on prevalence $h$ upon hearing utterances corresponding to the uncertain threshold (Eq. \ref{eq:L0}) as well as a comparison model using a fixed threshold at 0 (intuitively, corresponding to the quantifier "some").
The differences in the semantics are easily seen when considering a uniform prior distribution over $h$ (Figure\ \@ref(fig:simulations); top row).
The fixed-threshold model rules out only the lowest possible prevalence ($h=0$) and returns a posterior very similar to the prior.
The behavior of the generalization is distinct from a fixed-threshold model: The higher the prevalence $h$, the more likely it is after hearing the generalization.

The intuition behind the model's behavior can be gleaned by imagining what the listener model $L$ would believe given different thresholds, and averaging over those possibilities.
If the threshold $\theta$ is very high, only the highest prevalences $h$ would pass the threshold.
If the threshold $\theta$ is slightly lower but still high, the highest prevalences $h$ would still pass the threshold, as would some that are slightly lower.
As the threshold $\theta$ takes on lower and lower values, more and more $h$'s would surpass it.
When $\theta$ is very low, almost all $h$ are consistent with it (akin to fixed-threshold model).
The listener averages over these possibilities: $h$'s with higher values pass more thresholds and the result is that higher $h$'s are more likely *a posteriori*.

In this paper, we are interested in explaining the variable *endorsements* of generalizations (e.g., "Birds lay eggs" is good; "Birds are female" is not) based on some known or believed prevalence $h$ (e.g., only female birds lay eggs).
Following the Rational Speech Act (RSA) framework [@Goodman2016], we describe an **endorsement model** as a Bayesian speaker's decision of whether or not to say the generalization to the listener $L$ (Eq. \ref{eq:L0}).^[A more general version of this model can relax the assumption that the speaker has access to a specific value of prevalence $h$ that he is trying to communicate.
Rather, the speaker may have a belief distribution over $h$ for the category $k$, corresponding to the speaker's beliefs about the prevalence of the property in the category. 
In this situation, we would define the speaker's utility to be the expected value of the informativity, which integrates over her belief distribution:
$S(u \mid k) \propto \exp{(\lambda \cdot {\mathbb E}_{h\sim P_{k}} \ln{ \int_{\theta} L(h, \theta \mid u)} \diff \theta )}$
]

\begin{equation} 
S(u \mid h) \propto \exp{(\lambda \cdot \ln{ \int_{\theta} L_0(h, \theta \mid u)}  \diff \theta )}
\label{eq:S1}
\end{equation}

Since the speaker $S$ doesn't know the threshold $\theta$, he must integrate over the likely values that the listener will infer. 
<!-- In RSA, language production and comprehension are defined by recursive Bayesian models describing speaker and listener. -->
Additionaly, following RSA, the speaker model is assumed to be an approximately rational Bayesian agent with degree with rationality governed by parameter $\lambda$, which acts as a soft-max function.

@@

```{r simulations, fig.cap = "Prior distributions over the prevalence of a feature (Left; grey) and listener model posterior distributions upon hearing an utterance with an uncertain threshold (Middle; red) and a fixed threshold (Right; blue), for five different schematic prior distributions. The uncertain threshold (Middle; red) is hypothesized to correspond to utterances that convey generalizations, while the fixed threshold (Right; blue) is a control model. Shapes of the priors were chosen to intuitivly correspond to properties labeling the distributions.", fig.width = 9, fig.height=5}
all.simulations %>%
  mutate(Distribution = paste(Distribution, utterance, sep = "_")) %>%
  filter(Distribution != "Some", Distribution != "Most", Distribution != "Prior_generic") %>%
  mutate(Distribution = factor(Distribution, levels = 
                             c("Prior_some",  "Posterior_Generalization_generic", "Posterior_Generalization_some"),
                                labels = c("Prior/Silence",  "Generalization (Uncertain threshold)", "Fixed threshold")),
         PriorShape = factor(PriorShape, levels = c("uniform", "barks","laysEggs","carriesMalaria","isFemale"),
                             labels = c("uniform", "bark","lay eggs","carry malaria","are female")
                                  #"eatsFood","areSick", "uniformRare", 
                                  #"carriesMalaria"
                                  )
         ) %>%
  ggplot(., aes(x = value, #y = prob, 
                #lty = Distribution, 
                color = Distribution, fill = Distribution))+
    geom_density(#stat= 'identity', 
                aes(y = ..scaled..),
                 size = 0.6, alpha = 0.7, adjust = 1.5)+
    #geom_bar(stat = 'identity', position = position_dodge(), color = 'black')+
    theme_few() +
    #scale_color_solarized()+
    #scale_fill_solarized()+
    #scale_fill_manual(values = c("#636363", "#abdda4", "#2b83ba", "#d7191c"))+
    #scale_color_manual(values = c("#636363", "#abdda4", "#2b83ba", "#d7191c"))+
   scale_fill_manual(values = c("#636363", "#d7191c", "#2b83ba"))+
    scale_color_manual(values = c("#636363", "#d7191c", "#2b83ba"))+
    #scale_linetype_manual(values = c(3, 4, 2, 1))+
    scale_linetype_manual(values = c(3, 4, 1))+
    scale_x_continuous(breaks = c(0, 0.5, 1), limits= c(0, 1))+
    scale_y_continuous(breaks = c(0, 1), limits= c(0, 1))+
    xlab("Degree of probability h") +
    ylab("Scaled probability density") +
  facet_grid(PriorShape~Distribution, scales = 'free')+
  guides(fill = F, color = F)+
  theme(strip.text.y = element_text(angle = 0))
```
